`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'` 


# The First Appendix

<!--
If you feel it necessary to include an appendix, it goes here.
-->

## Ethics information

All procedures for this project have been approved by the University of Calgary Conjoint Faculties Research Ethics Board (REB 20-1860). For the experimental studies, informed consent will be obtained for all participants, who, unless otherwise noted, will be compensated at a rate of \$12.48 CAD/hour. For multilevel analyses, the datasets are archival and publicly-available. As such, ethics approval and informed consent was obtained, and compensation disbursed, by the original researchers.

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.

**In the main Rmd file**

```{r ref.label='include_packages', results='hide', echo = F}
```

**In Chapter \@ref(ineq-fair):**

```{r ref.label='include_packages_2', results='hide', echo = F}
```

# The Second Appendix, for Fun
