---
title: 'Examining the Role of Inequality in Security Consumption'
author: 'Dallas R. Novakowski'
date: 'AAAA 2022'
month: 'xxxxx'
year: '2022'
institution: 'UNIVERSITY OF CALGARY'
division: 'HASKAYNE SCHOOL OF BUSINESS'
advisor: 'Mehdi Mourali'
# If you have more two advisors, un-silence line 7
#altadvisor: 'Your Other Advisor'
department: 'Marketing'
degree: 'DOCTOR OF PHILOSOPHY'
location: 'CALGARY, ALBERTA'
knit: bookdown::render_book
site: bookdown::bookdown_site
link-citations: yes


# The next two lines allow you to change the spacing in your thesis. You can 
# switch out \onehalfspacing with \singlespacing or \doublespacing, if desired.
header-includes:
    - \usepackage{setspace}\onehalfspacing

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True
  
# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
  thesisdown::thesis_pdf: default 
#  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  Mehdi, Scott Radford, James, Bierman, John Van Kesteren, Sharon, 
  philip, Zhanna, Quan, Kerry
  Open science communities - R, Otree 
dedication: |
  Blah blah
preface: |
  This thesis is original, unpublished, independent work by the author, D. Novakowski
  
# Specify the location of the bibliography below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```


```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```


<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

Income and wealth inequalities have risen sharply in the US in the past 30 years [@horowitz2020] Inequality's moral and political significance is closely linked to its impacts on people's psychology and behaviour. At a societal level, inequality has been associated with higher nation-level homicide rates [@daly2016], consumer debt [@christen2005], and reduced cooperation [@burton-chellew2013 ; @anderson2008], while at an individual level, inequality causes people to be more likely to resort to risky, hostile, and criminal activities in desperate attempts to compete for otherwise inaccessible positions and resources [@mishra2015; @payne2017; @krupp2018]. This dissertation seeks to examine whether inequality causes a special case of risk aversion: social distrust, and increases individuals' willingness to consume security products [@payne2017].

Chapter 1 will provide an initial test of the inequality-distrust relationship through a series of economic games, using crowdsourced participants. These experiments will manipulate inequality and test a mechanism of envy avoidance: that inequality increases distrust and security consumption because participants expect their partners to be frustrated by a disparity in incomes. Using the same game context with crowdsourced participants, Chapter 2 examines the impact of distributional fairness on security consumption, manipulating whether incomes are deserved, and whether incomes are allocated as a product of group-based discrimination. Chapter 3 describes a multilevel analysis using the International Crime Victimization Survey, testing whether the country-level income inequality is associated with individual-level security consumption.

## Inequality and Risk-Taking

Two accounts have been proposed to explain how population-level inequality influences risk-taking in individuals. Both of which follow from risk-sensitivity theory, which predicts that organisms will take risks when more predictable strategies cannot meet the organism's goals [e.g., starvation thresholds; see @caraco1980]. Accordingly, inequality fosters risk-taking by increasing the distance between individuals and their goals. First, inequality increases risk-taking through *absolute deprivation*; unequal societies tend to have more poverty, and poor people are more likely to take risks to meet their basic needs [@pridemore2008]. Second, risk-taking can be a response to envy (or *relative deprivation*), where even well-off individuals feel disadvantaged when compared to others, then take risks in an attempt to keep up with peers and a socially-defined reference point [@mishra2017].

The relative state model [@mishra2017] posits that individuals take risks not only to escape material destitution, but also to keep up with competitors. As a result, risk-taking can be influenced by social comparisons. As wealth distributions become more skewed, so do comparison standards. Under conditions of inequality, individuals set more ambitious goals, and subsequently take more risks, as they try to close the gap between their current state and their aspiration levels[@payne2017]. Likewise, being worse off than others can lead to emotional responses of envy and personal relative deprivation, which can be linked to gambling as well as spiteful behaviours [@callan2008; @wobker2015]. 

This dissertation is inspired by considering the social consequences of this inequality-risk relationship. If inequality causes people to take more risks and/or behave in a hostile manner, would people in the same unequal environment come to expect harmful behaviours from their neighbors? Chapter one will elaborate on this envy aversion hypothesis of distrust, which predicts that in the presence of inequality, concerns over partners' envy will drive people to pay to secure themselves against social threats.  

## Security and Trust - A Typology of Self-Protection

The pursuit of security is fundamental to human life; we lock our car doors, purchase insurance, stockpile protective equipment, and flee unstable countries. Many markets are directly and indirectly influenced by this desire for security. In the United States, insurance premiums totaled to \$1.2 trillion in 2017 [@insuranceinformationinstitute2020a]. Likewise, people who are concerned for their safety are less likely to use of public transportation [@delbosc2012; @loukaitou-sideris1997]. The pursuit of security can require large tradeoffs for consumers; every dollar spent on a heavy-duty deadbolt or theft insurance cannot be spent on other goods such as savings, recreation, or healthcare. Individuals' purchasing and use of products that seek to increase one's security are of particular interest for this project, or what will be termed security consumption.

The idea of security is broad, rooted in the Latin term *securitas*, meaning freedom from care or anxiety. So, (in)security can occur in many areas, ranging from health, employment, natural disasters, nutrition, as well as conflicts between individuals or nations. Given its potential breadth, a concept of security consumption must be clarified in terms of its relevant domains and forms. At its most comprehensive, security consumption could be considered as any prevention-focused activity, conducted by an agent seeking to minimize the expected costs from harms. With such a broad definition, security consumption can be carried out by individual consumers, organizations, as well as nation-states.

Scholars in risk management classify any source of costs as a hazard, whereas threats are defined as warnings that something unpleasant may occur [@brunschot2007]. Given the conceptual breadth of consumer behaviour [@macinnis2010], security consumption can be defined as an expenditure of resources in the consumer context that helps the agent to (a) recover from, (b) cope with, (c) avoid, and/or (d) deter hazards. Notably, security behaviours do not always manifest in the consumer context; for instance, someone might be concerned about being victimized by a burglary and develop habits of checking their locks, or limiting their exposure to high-risk areas. Although such vigilance might improve one's ability to cope with an intruder, or prevent victimization altogether, it does not manifest in meaningful exchanges within the security market. Thus, security behaviours are best understood as antecedent or moderating variables in the decision to consume security goods and services.

*Recovery* corresponds namely to purchasing insurance. For instance, although automobile insurance is necessary for all drivers, some people purchase extra insurance to help recuperate costs from collisions or vandalism. *Coping* refers to goods and services that help agents reduce the costs incurred when exposed to harms. For instance, a fire extinguisher helps in coping with flammable hazards, such that when it occurs, a fire can be smothered before it causes substantial harm. Likewise, individuals can expend resources to *avoid* certain hazards, for instance, people might invest substantial resources in the process of moving away from regions that suffer from crime, air pollution, or natural disasters. Lastly, agents can consume security goods that are meant as *deterrents* for potentially antagonistic actors. For instance, someone might purchase a "beware of dog" sign for their property to deter intruders, when they in fact do not own a dog.

Importantly, many security goods and services fill multiple roles simultaneously. For example, moving into a gated community may 1) help their residence avoid being near and visible to criminals, 2) have protective measures that help to cope with attempted thefts, and 3) reduce criminals' willingness to attempt a theft, due to the perceived risk of getting caught or hurt. This definition of security consumption appears to be capable of incorporating and differentiating a wide variety of security behaviours.

The physical security industry has been projected to increase 29% between 2021 and 2025 [@marketsandmarkets2021], which likely is heavily impacted by organizations and nations. Understanding security consumption by institutions (i.e., business-to-business and business-to-government marketing) is important, as government and private organizations likewise seek to recover from, cope with, avoid, and deter harms. However, the contexts of such institutional decisions (e.g., institutional prestige, political leverage), the hazards (nuclear strikes by other countries), and the products available (armed drones), are distinct from the security decisions of everyday consumers. To avoid confusing organizations and individuals, this proposal will focus on security consumption amongst individual consumers, such limitation in scope will help to develop and explicate microfoundations of security consumption, which may eventually be generalized or modified to other security market contexts.

## Social Conflict as a Security Hazard

As mentioned earlier, hazards can come from a variety of sources (natural disasters, economic instability, pandemics). However, this proposal will focus on the threats posed by other people, and consequently, the consumption of deterrence-focused goods. This emphasis on social sources of harm reveals how decision-makers must consider others' intentions when deciding whether to purchase security goods. Specifically, security behaviours are heavily influenced by the hypothetical imperative: *if I expect others to [try to steal]{.ul}, my best response is to [purchase security products]{.ul}*; [@hargreaves-heap2004a]. Correspondingly, this proposal will examine how these beliefs of others' intentions and of the larger social environment might influence the consumption of security goods.

Ensuring the safety of one's body and resources against hostile behaviours is an essential life function. For instance, tadpoles are quick to abandon rich food sources when exposed to a chemical "alarm substance" secreted from other injured tadpoles [@petranka1989]. The risk of predation has been argued to have selected for vigilance and defensive strategies across the animal kingdom [e.g., @beauchamp2015]. The prevalence of defensive strategies can be seen in the tissue-regeneration of starfishes, aggressive behaviours in honey badgers, camouflage in stick insects, and the bright coloration of non-venomous animals such as the milk snake. Likewise, humans and our ancestors have historically faced predation by other animals [@hart2011], which have been argued to yield risk-monitoring adaptations in modern humans, such as collective risk-monitoring [@kameda2007].

While the risk of predation by other species seems to have a profound effect on security behaviours in humans, otherpeoplehave proven to be a substantial hazard to our ancestors and contemporaries alike. Currently, other humans (via murder) are only outranked by mosquitoes as the deadliest animal to modern *Homo sapiens* [@gates2014]. Given the costs of being victim to acts of hostility, and the benefits of cooperation [@henrich2001a], the identification of friends versus enemies is a fundamental strategic problem in human interaction. Given humans' characteristic reliance on markets to meet their needs, the remainder of this proposal will rest on the assumption that security consumption (that protect against social hazards) is at least in part an expression of consumer expectations of antagonism from actors in their social environment.


-   Summarize each chapter (i.e., paper) that has been included in your thesis
-   State the general theme(s) of your research and the papers you've included
-   Include a Statement of Contribution that outlines the work you and your co- authors contributed to each paper with multiple authors
-   May include literature relevant to your topic that doesn't appear in the chapt

